---
title: "Ecology Assignment 2"
author: "Sandra Fogg"
date: "3/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Load Packages in Console

#Load Data Sets
fertilizer_df <- read_csv("fertilizer.csv")
grain_df <- read_csv("grain.csv")


```
##1. Patterns in Yield Over Time

####Create a Finalized Graph of Yield Over Time for Each Grain
```{r}
#Create All Crops Dataframe
all_crops_clean <- grain_df %>%
  filter(yield != "NA")
#Create All Crops Plot
all_crops_plot_yield <- ggplot(all_crops_clean, aes(x=year, y=yield)) +
  geom_point(aes(color = commodity), alpha=0.4) +
  labs(y = "Crop Yield (bushels/acre)", x = "Year", color = "Commodity") +
  theme_classic() +
  ggtitle("US Agricultural Crop Yield from 1866 to 2016") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 180), breaks = seq(0, 180, 20)) +
  scale_x_continuous(expand= c(0,0), breaks = seq(1866, 2016, 20))
  
all_crops_plot_yield

all_crops_animated <- ggplotly(all_crops_plot_yield)
all_crops_animated
all_crops_motion <- gganimate(all_crops_plot_yield)


#SAME PLOT BUT FACET WRAPPED FOR SEPARATENESS
all_crops_plot_FACET <- ggplot(all_crops_clean, aes(x=year, y=yield)) +
  geom_line() +
  labs(y = "Crop Yield (bushels/acre)", x = "Year", color = "Commodity") +
  theme_classic() +
  ggtitle("US Agricultural Crop Yield from 1866 to 2016") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 180), breaks = seq(0, 180, 30)) +
  scale_x_continuous(expand= c(0,0), breaks = seq(1866, 2016, 30))+
  facet_wrap(~commodity, scale = "free")
  
all_crops_plot_FACET


#SEPARATE PLOT FOR EACH GRAIN
#Plot for Corn
corn_yield_clean <- all_crops_clean %>% 
  filter(commodity == "Corn")
corn_yield_plot <- ggplot(corn_yield_clean, aes(x=year, y=yield)) +
  geom_line() +
  labs(y = "Crop Yield (bushels/acre)", x = "Year") +
  theme_classic() +
  ggtitle("US Corn Yield from 1866 to 2016") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 180), breaks = seq(0, 180, 20)) +
  scale_x_continuous(expand= c(0,0), breaks = seq(1866, 2016, 20))
corn_yield_plot

#Plot for Barley
barley_yield_clean <- all_crops_clean %>% 
  filter(commodity == "Barley")
barley_yield_plot <- ggplot(barley_yield_clean, aes(x=year, y=yield)) +
  geom_line() +
  labs(y = "Crop Yield (bushels/acre)", x = "Year") +
  theme_classic() +
  ggtitle("US Barley Yield from 1866 to 2016") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 100), breaks = seq(0, 100, 20)) +
  scale_x_continuous(expand= c(0,0), breaks = seq(1866, 2016, 20))
barley_yield_plot

#Plot for Oats
oats_yield_clean <- all_crops_clean %>% 
  filter(commodity == "Oats")
oats_yield_plot <- ggplot(oats_yield_clean, aes(x=year, y=yield)) +
  geom_line() +
  labs(y = "Crop Yield (bushels/acre)", x = "Year") +
  theme_classic() +
  ggtitle("US Oats Yield from 1866 to 2016") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 100), breaks = seq(0, 100, 20)) +
    scale_x_continuous(expand= c(0,0), breaks = seq(1866, 2016, 20))
oats_yield_plot


#Plot for Sorghum
sorghum_yield_clean <- all_crops_clean %>% 
  filter(commodity == "Sorghum")
sorghum_yield_plot <- ggplot(sorghum_yield_clean, aes(x=year, y=yield)) +
  geom_line() +
  labs(y = "Crop Yield (bushels/acre)", x = "Year") +
  theme_classic() +
  ggtitle("US Sorghum Yield from 1919 to 2016") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 100), breaks = seq(0, 100, 20)) +
    scale_x_continuous(expand= c(0,0), breaks = seq(1910, 2016, 10))
sorghum_yield_plot


```
Corn is linear with a lower plateau.
Barley is linear with a lower plateau.

##2. Fertilizer Use

####Create a Finalized Plot of Fertilizer Use vs. Time for Each Crop
```{r}
#Create Wheat Dataframe
fertilizer_wheat <- fertilizer_df %>%
  filter(ppa != "NA",
         crop == "Wheat")
#Create Wheat Plot
wheat_plot <- ggplot(fertilizer_wheat, aes(x=year, y=ppa)) +
  geom_line(aes(color = fertilizer)) +
  labs(y = "Fertilizer Application (lbs/acre)", x = "Year", color = "Fertilizer") +
  theme_classic() +
  ggtitle("US Agricultural Fertilizer Use on Wheat from 1964 to 2016") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 100), breaks = seq(0, 100, 10)) +
  scale_x_continuous(expand= c(0,0), breaks = seq(1964, 2024, 10))+
  scale_color_brewer(palette = "Dark2")
  
wheat_plot


#Create Corn Dataframe
fertilizer_corn <- fertilizer_df %>%
  filter(ppa != "NA",
         crop == "Corn")
#Create Corn Plot
wheat_plot <- ggplot(fertilizer_corn, aes(x=year, y=ppa)) +
  geom_line(aes(color = fertilizer)) +
  labs(y = "Fertilizer Application (lbs/acre)", x = "Year", color = "Fertilizer") +
  theme_classic() +
  ggtitle("US Agricultural Fertilizer Use on Corn from 1964 to 2016") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 100), breaks = seq(0, 100, 10)) +
  scale_x_continuous(expand= c(0,0), breaks = seq(1964, 2024, 10))+
  scale_color_brewer(palette = "Dark2")
  
Corn_plot


#Create Soybean Dataframe
fertilizer_soybean <- fertilizer_df %>%
  filter(ppa != "NA",
         crop == "Soybeans")
#Create Soybean Plot
soybean_plot <- ggplot(fertilizer_soybean, aes(x=year, y=ppa)) +
  geom_line(aes(color = fertilizer)) +
  labs(y = "Fertilizer Application (lbs/acre)", x = "Year", color = "Fertilizer") +
  theme_classic() +
  ggtitle("US Agricultural Fertilizer Use on Soybeans from 1964 to 2016") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 100), breaks = seq(0, 100, 10)) +
  scale_x_continuous(expand= c(0,0), breaks = seq(1964, 2024, 10))+
  scale_color_brewer(palette = "Dark2")
  
soybean_plot


#Create All Crops Plot
all_crops_plot <- ggplot(fertilizer_df, aes(x=year, y=ppa)) +
  geom_point(aes(color = fertilizer), alpha = 0.5) +
  labs(y = "Fertilizer Application (lbs/acre)", x = "Year", color = "Fertilizer") +
  theme_classic() +
  ggtitle("US Agricultural Fertilizer Use from 1964 to 2016") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 150), breaks = seq(0, 150, 20)) +
  scale_x_continuous(expand= c(0,0), limits = c(1964, 2018), breaks = seq(1964, 2018, 10))+
  scale_color_brewer(palette = "Dark2") +
  facet_wrap(~crop, scale = "free") 
  
all_crops_plot
```
####Describe Differences in Fertilizer Use for Each Crop

Corn and Wheat are primarily treated with Nitrogen fertilizer, although corn requires significantly more fertilizer than wheat, suggesting that Nitrogen is the limiting factor for plant growth. Corn's extensive nutrient needs may be attributable to an inefficient uptake of Nitrogen.  Soybeans are instead predominantly treated with Potassium, and are treated with very small amounts of Nitrogen.  This is probably due to the nitrogen-fixing symbiotic organisms that live in the nodules of soybean plants, which enable them to acquire more Nitrogen from the surrounding soils.  It is thus likely that Potassium is the limiting nutrient for soybean growth. While Phosphate fertilizer levels are a bit higher for corn, the three crops don't appear to have a drastic difference in Phosphate application and require similar levels.  Also it is important to note that the ratios of N:P:K vary a lot between the three crops.  Corn and soybeans appear to have much larger ratios of the three nutrients, while the demand for nutrients is more evenly distributed for wheat.  


##3. Case Study: Corn and Fertilizer

####Create a Finalized Plot of Corn Yield vs. PPA
```{r}
#Create Corn Yield Dataframe
grain_years_df <- grain_df[grain_df$year >= 1963 & grain_df$year <=2017,] 
corn_yield <- grain_years_df %>% 
  filter(yield != "NA",
         commodity == "Corn")

corn_yield_use <- merge(fertilizer_corn,corn_yield, sort = TRUE)
corn_yield_use
#Create Corn Yield Plot
corn_yield_fert_plot <- ggplot(corn_yield_use, aes(x = ppa, y = yield)) +
  geom_point(aes(color = year, shape = fertilizer), size = 3, alpha = 0.9) +
  labs(y = "Crop Yield (bushels/acre)", x = "Fertilizer Application (lbs/acre)", color = "Year", shape = "Fertilizer") +
  theme_classic() +
  scale_color_gradient(low = "firebrick2", high = "chartreuse") +
  ggtitle("US Fertilizer Use on Corn and Resulting Crop Yield from 1964 to 2016") +
  scale_y_continuous(expand = c(0,0), limits = c(30, 180), breaks = seq(0, 180, 20)) +
  scale_x_continuous(expand= c(0,0), limits = c(30, 160), breaks = seq(0, 160, 20))

  
corn_yield_fert_plot
 
```
Similar to the plot shown in class about cereal yield and Nitrogen fertilizer use in countries of varying stages in development, there appears to be positive linear relationship between the two.  Intensification of Nitrogen fertilizer has increased by almost four times the amount applied in the 1960s, producing higher crop yields as a result. However, the other two fertilizers, Phosphate and Potassium, do not seem to share this trend.  While it appears there have been small increases in Phosphate and Potassium fertilizers since the 1960s, application amounts have appeared to remain relatively steady.  Despite the consistency of fertilizer application, yields have increased over time.  It is possible that this is due to more efficient strategies of Phosphate and Potassium fertilizer delivery.  
